Curve counting in genus one: \\ elliptic singularities {\it \&} relative geometry

### 2021-03-28 Response to referee/AG Genus One Response to Referee.tex

\def\bibfont{\footnotesize}

\begin{document}

\title{Response to referee report: AG 608 (Version 1)}
%\author{}
\maketitle

\noindent We thank the referee for their careful reading of our paper, and the many valuable suggestions which have greatly helped in improving the quality of the exposition.

We have made many improvements to the paper in order to directly address the referee's comments, without exceptions. Following their suggestion, we have also gone through and made additional changes, in order to improve the readability and appeal to a wider audience. While adding more exposition, we have strived to cut down on unnecessary words and phrases and streamline the language, so the length of the paper has not changed substantially. Major such changes include:
\begin{enumerate}
	\item The introduction has been streamlined to better reflect the growing body of work on the connections between curve singularities, logarithmic structures and stable maps.
	\item An improved background section (see \S\S 1.1--1.5), explicating the connection between logarithmic structures and moduli of Gorenstein singularities, which is a key geometric ingredient in the paper. In particular, we have included working definitions of the background material. 
	\item The argument proving Theorem 1.9 has been reworked to have a clearer logical structure, with additional details. 
	\item The arguments of \S 4.3 have been expanded and streamlined.
\end{enumerate}

@@

\item[(31)] We point out that this uses the same arguments as in Step~1. In the Step~1 arguments, we add explicit references to Gathmann's genus zero results and the necessary recursions on genus one Deligne--Mumford space.
\item[(32)] We have broken up Step 2 and Step 4 into smaller pieces. At several points we have made use of numbered lists for case analysis, which clarifies the logical structure. We have also provided summaries within steps as well as at the end of each step, to help the reader follow which parts of the formula the present arguments pertain to.
\end{enumerate}

%\subsection*{Additional minor changes}
%\begin{enumerate}
%\item The term ``special rubber'' used at the start of Step 2 is an anachronism from a draft version of the paper, and has been removed.
%\end{enumerate}

### 2021-03-28 EllipticSingularitiesSplitting.tex
Imported Dhruv v2 submmited to AG

The higher genus situation is more delicate, and substantial theory has been developed in recent years to treat the genus one case, both for its intrinsic geometry, and for higher genus insights. This began over a decade ago with pioneering work of Li, Vakil, and Zinger~\cite{redgone,VZ,LZ,lz2,zingerstvsred,zingred,HLN}, and was recently reinvigorated using ideas from singularity theory~\cite{BCM18,HL,SMY1,VISC} and logarithmic geometry~\cite{RSPW,RSPW2,BC20}. The result is a \textbf{reduced} Gromov--Witten theory, which removes degenerate contributions from the ordinary theory. Fundamental ingredients in Gromov--Witten theory are torus localisation, relative stable maps, and degeneration formulas. While the Atiyah--Bott localisation for the reduced invariants was used to great effect by Zinger in his proof of the BCOV conjectures, the remaining components -- relative invariants and the degeneration formula -- have not been developed in this setting. Our work provides these as new {structural} tools in reduced genus one Gromov--Witten theory\footnote{Further motivation to consider the genus one case specifically comes simply from the fact that for Calabi--Yau geometries in dimension larger than $3$, Gromov--Witten theory vanishes in genera greater than one.}.

@@

\subsection{Context and techniques} The present work is part of a larger ongoing program relating to the role of curve singularities in Gromov--Witten theory that go beyond the well-studied nodal case~\cite{BC20,BCM18,Boz19,RSPW,RSPW2}. The basic motivation for this direction is that such alternative curve counting theories can often be more {enumerative} in nature and satisfy Lefschetz section theorems. Degeneration methods are a crucial tool in Gromov--Witten theory, and our results point to a fruitful use of these methods in this direction.

@@

We put the pieces together. The space $ \operatorname{Obs}(f/(C,\delta_B,f_B))$ always has rank $1$. Next, note that once the collapsed stable map is fixed, there are no nontrivial \textit{logarithmic} deformations of the map to $\mathbb P^m[s]$: the morphism from logarithmic maps to ordinary maps is finite~\cite[\S~3.7]{ChenLog}. Therefore, we may view the logarithmic deformation spaces of $f$ as a subspace of the logarithmic deformation spaces of $(C^\prime,\delta_B,f_B^\prime)$, consisting of those deformations with vanishing class in $\operatorname{Obs}(f/(C,\delta_B,f_B))$. The data $(C^\prime,\delta_B,f_B^\prime)$ have unobstructed deformations, so this subspace is everywhere of the expected dimension. The obstructions to logarithmic deformations of $f$ vanish, and we conclude.
\end{proof}

@@

\begin{proof}
Since $\VZ_{1,\alpha}(\mathbb P^m|H,d)$ is logarithmically smooth, it will suffice to show that the cones of its tropicalization are simplicial. Consider a logarithmic stable map to an expansion $C\to \mathbb P^m[s]$ without a central alignment. The tropical moduli cone obtained as the dual of the minimal base monoid can be identified with $\mathbb R_{\geq 0}^{s}$, see for instance~\cite[Section~2.2]{ChenDegeneration}. The alignment is an iterated stellar subdivision on tropical moduli spaces, see~\cite[Section 4.6]{RSPW}, and such subdivisions preserve the property of being simplicial. We conclude that the blowup $\widetilde{\VZ}^{\mathrm{exp}}_{1,\alpha}(\mathbb P^m|H,d)$ obtained by centrally aligning Kim's spaces has simplicial cones. The morphism $\VZ_{1,\alpha}(\mathbb P^m|H,d)\to \widetilde{\VZ}^{\mathrm{exp}}_{1,\alpha}(\mathbb P^m|H,d)$ is strict, so the cones of $\VZ_{1,\alpha}(\mathbb P^m|H,d)$ are simplicial, as claimed.
\end{proof}

@@

\begin{proof}
We first consider the statement of the lemma in the category of fine but not necessarily saturated logarithmic schemes, so that a relative stable map to an expansion admits a unique logarithmic lifting. The fibre product description preceding the lemma is an isomorphism in this category~\cite[Lemma~4.2.2]{AbramovichMarcusWiseComparison}.

@@

\begin{remark}[Splitting degree and vanishing order, revisited] Similarly to \S \ref{subsubsection splitting degree} above, we can calculate the degree of $\rho$ and the vanishing order $\lambda_{\Dcal}$ of the section $s_k$ along $\Dcal$. The degree of $\rho$ is given exactly as in Lemma \ref{lem:saturation}, where now we obtain one such factor for each bounded edge of the tropical target. The vanishing order $\lambda_{\Dcal}$ is slightly more combinatorially delicate, but may be computed as the index of an explicit morphism of one-dimensional lattices, as explained in the proof of Lemma \ref{lemma vanishing order}.
\end{remark}

>>>>

The higher genus situation is more delicate, and substantial theory has been developed in recent years to treat the genus one case, both for its intrinsic geometry, and for higher genus insights. This began over a decade ago with pioneering work of Li, Vakil, and Zinger~\cite{redgone,VZ,LZ,lz2,zingerstvsred,zingred}, and was recently reinvigorated using ideas from singularity theory~\cite{BCM18,HL,SMY1,VISC} and logarithmic geometry~\cite{RSPW,RSPW2,BC20}. The result is a \textbf{reduced} Gromov--Witten theory, which removes degenerate contributions from the ordinary theory. Fundamental ingredients in Gromov--Witten theory are torus localisation, relative stable maps, and degeneration formulas. While the Atiyah--Bott localisation for the reduced invariants was used to great effect by Zinger in his proof of the BCOV conjectures, the remaining components -- relative invariants and the degeneration formula -- have not been developed in this setting. Our work provides these as new {structural} tools in reduced genus one Gromov--Witten theory\footnote{Further motivation to consider the genus one case specifically comes simply from the fact that for Calabi--Yau geometries in dimension larger than $3$, Gromov--Witten theory vanishes in genera greater than one.}.

@@

\subsection{Context and techniques} The present work is part of a larger ongoing program relating to the role of curve singularities and alternate compactifications in Gromov--Witten theory that go beyond the well-studied geometry of stable maps from nodal curves~\cite{BC20,BCM18,Boz19,HLN,RSPW,RSPW2}. The motivation for this direction is that such alternative curve counting theories can often be more {enumerative} in nature and satisfy Lefschetz section theorems. Degeneration methods are a crucial tool in Gromov--Witten theory, and our results point to a fruitful use of these methods in this direction.

@@

We put the pieces together. The space $ \operatorname{Obs}(f/(C,\delta_B,f_B))$ always has rank $1$. Next, note that once the collapsed stable map is fixed, there are no \textit{logarithmic} deformations of the map to $\mathbb P^m[s]$: the morphism from logarithmic maps to ordinary maps is finite~\cite[\S~3.7]{ChenLog}. View the logarithmic deformation spaces of $f$ as subspace of the logarithmic deformation spaces of $(C^\prime,\delta_B,f_B^\prime)$, consisting of those deformations with vanishing class in $\operatorname{Obs}(f/(C,\delta_B,f_B))$. The map to this obstruction group is everywhere surjective, and since the data $(C^\prime,\delta_B,f_B^\prime)$ have unobstructed deformations, this subspace is everywhere of the expected dimension. Obstructions to logarithmic deformations of $f$ vanish, and we conclude.
\end{proof}

@@

\begin{proof}
Since $\VZ_{1,\alpha}(\mathbb P^m|H,d)$ is logarithmically smooth, we must show that the cones of its tropicalization are simplicial. Consider a logarithmic stable map to an expansion $C\to \mathbb P^m[s]$ without a central alignment. The tropical moduli cone obtained as the dual of the minimal base monoid can be identified with $\mathbb R_{\geq 0}^{s}$, see for instance~\cite[Section~2.2]{ChenDegeneration}. The alignment is an iterated stellar subdivision on tropical moduli spaces, see~\cite[Section 4.6]{RSPW}, and such subdivisions preserve the property of being simplicial. We conclude that the blowup $\widetilde{\VZ}^{\mathrm{exp}}_{1,\alpha}(\mathbb P^m|H,d)$ obtained by centrally aligning Kim's spaces has simplicial cones. The morphism $\VZ_{1,\alpha}(\mathbb P^m|H,d)\to \widetilde{\VZ}^{\mathrm{exp}}_{1,\alpha}(\mathbb P^m|H,d)$ is strict, so the cones of $\VZ_{1,\alpha}(\mathbb P^m|H,d)$ are simplicial.
\end{proof}

@@

\begin{proof}
We consider the statement of the lemma in logarithmic schemes, so that a relative stable map to an expansion admits a unique logarithmic lifting, and later saturate the structures. The fibre product description preceding the lemma is an isomorphism in this category~\cite[Lemma~4.2.2]{AbramovichMarcusWiseComparison}.

@@

\begin{remark}[Splitting degree and vanishing order, revisited] Similarly to \S \ref{subsubsection splitting degree} above, we can calculate the degree of $\rho$ and the vanishing order $\lambda_{\Dcal}$ of the section $s_k$ along $\Dcal$. The degree of $\rho$ is given exactly as in Lemma \ref{lem:saturation}, where we obtain one such factor for each bounded edge of the tropical target. The vanishing order $\lambda_{\Dcal}$ is more delicate, but may be computed as the index of an explicit morphism of one-dimensional lattices, explained in the proof of Lemma \ref{lemma vanishing order}.
\end{remark}
